Model Checking Differentially Private Properties

### 2018-06-14 semantics.tex
add Depeng and Lijun's fix

Other than the differentially private operator, the semantics of
$\dpCTLstar$ is standard~\cite{BK:08:PMC}.
To intuit the semantics of $\dpriv{\epsilon}{\delta} \phi$,
recall that  $\myPr{s}{K}{N}{\phi}$ is the probability of having
paths satisfying $\phi$ from $s$. A state $s$ satisfies
$\dpriv{\epsilon}{\delta} \phi$ if the probability of
having paths satisfying $\phi$ from $s$ is $(\epsilon, \delta)$-close
those from every neighbor of $s$. Informally, it is
probabilistically similar to observe paths satisfying $\phi$ from $s$
and from its neighbors.

>>>>

Other than the differentially private operator, the semantics of
$\dpCTLstar$ is standard~\cite{BK:08:PMC}.
To intuit the semantics of $\dpriv{\epsilon}{\delta} \phi$,
recall that  $\myPr{s}{K}{N}{\phi}$ is the probability of having
paths satisfying $\phi$ from $s$. A state $s$ satisfies
$\dpriv{\epsilon}{\delta} \phi$ if the probability of
having paths satisfying $\phi$ from $s$ is $(\epsilon, \delta)$-close
to those from every neighbor of $s$. Informally, it is
probabilistically similar to observe paths satisfying $\phi$ from $s$
and from its neighbors.

### 2018-06-14 specifications.tex
add Depeng and Lijun's fix

For the $\frac{1}{2}$-geometric mechanism (Section~\ref{subsec:geometric}),
define the formula
$\psi = \dpriv{\ln 2}{0} (\X out_0) \wedge \dpriv{\ln 2}{0} (\X out_1) \wedge
\cdots \wedge \dpriv{\ln 2}{0} (\X out_5)$. If the state $s_k$ satisfies
$\psi$ for $k = 0, \ldots, 5$, then the $\frac{1}{2}$-geometric mechanism is
$(\ln 2, 0)$-differentially private. For the subsampling majority
mechanism (Section~\ref{subsection:subsampling}), consider the formula
$\psi = \dpriv{\epsilon, \delta} (\F 0) \wedge \dpriv{\epsilon,
  \delta} (\F 1)$. If a state satisfies $\psi$, its probability of
outputting is $(\epsilon, \delta)$-close to those of its
neighbor for every outcomes. The subsampling majority
mechanism is $(\epsilon, \delta)$-differentially private.

@@

\paragraph{Tighter Privacy Bounds for Composition.}
An advantage of applying model checking is that we may get tighter
bounds for composition. Consider the survey mechanism, and the
property $\dpriv{0}{.5} (\X \mathit{out}_1)$. Obviously, it holds in
states $+$ and $-$ since $\myPr{u}{K}{N}{\mathit{out}_1} =
\frac{3}{4}$, $\frac{1}{4}$ for $u = +, -$ respectively (Figure~\ref{figure:2-dp}). A careful check infers that one cannot decrease  $\delta_1=.5$ without increasing $\epsilon$.
Now consider the formula $\dpriv{\epsilon_2}{\delta_2} (\X
(\mathit{out}_1\wedge \X \mathit{out}_1))$ in
Figure~\ref{figure:2-dp-mc}. Applying the
compositional theorem, one has $\epsilon_2=2\epsilon_1=0$ and
$\delta_2=2\delta_1=1$. However, we can check easily that one gets
better privacy parameter $(0,.5)$ using the model checking algorithm
because $\myPr{u}{K}{N}{\mathit{out}_1} = \frac{9}{16}$,
$\frac{1}{16}$ for $u = +, -$ respectively.
In general, compositional theorems for differential privacy only give
asymptotic upper bounds. Privacy parameters $\epsilon$ and $\delta$
must be calculated carefully and often pessimistically. Our
algorithm allows data analysts to choose better parameters.

>>>>

For the $\frac{1}{2}$-geometric mechanism (Section~\ref{subsec:geometric}),
define the formula
$\psi = \dpriv{\ln 2}{0} (\X out_0) \wedge \dpriv{\ln 2}{0} (\X out_1) \wedge
\cdots \wedge \dpriv{\ln 2}{0} (\X out_5)$. If the state $s_k$ satisfies
$\psi$ for $k = 0, \ldots, 5$, then the $\frac{1}{2}$-geometric mechanism is
$(\ln 2, 0)$-differentially private. For the subsampling majority
mechanism (Section~\ref{subsection:subsampling}), consider the formula
$\psi = \dpriv{\epsilon}{\delta} (\F 0) \wedge
\dpriv{\epsilon}{\delta} (\F 1)$. If a state satisfies $\psi$, its
probability of 
outputting is $(\epsilon, \delta)$-close to those of its
neighbor for every outcomes. The subsampling majority
mechanism is $(\epsilon, \delta)$-differentially private.

@@

\paragraph{Tighter Privacy Bounds for Composition.}
An advantage of applying model checking is that we may get tighter
bounds for composition. Consider the survey mechanism, and the
property $\dpriv{0}{.5} (\X \mathit{out}_1)$. Obviously, it holds in
states $+$ and $-$ since $\myPr{u}{K}{N}{\mathit{out}_1} =
\frac{3}{4}$, $\frac{1}{4}$ for $u = +, -$ respectively (Figure~\ref{figure:2-dp}). A careful check infers that one cannot decrease  $\delta_1=.5$ without increasing $\epsilon$.
Now consider the formula $\dpriv{\epsilon_2}{\delta_2} (\X
(\mathit{out}_1\wedge \X \mathit{out}_1))$ in
Figure~\ref{figure:2-dp-mc}. Applying the
compositional theorem, one has $\epsilon_2=2\epsilon_1=0$ and
$\delta_2=2\delta_1=1$. However, we can get
better privacy parameter $(0,.5)$ using the model checking algorithm
since $\myPr{u}{K}{N}{\X (\mathit{out}_1 \wedge \X\mathit{out}_1)}
= \frac{9}{16}$, $\frac{1}{16}$ for $u = +, -$ respectively.
Generally, compositional theorems in differential privacy only give
asymptotic upper bounds. Privacy parameters $\epsilon$ and $\delta$
must be calculated carefully and often pessimistically. Our
algorithm allows data analysts to choose better parameters.
